\LARGE Coursework \\ Module Name (CMM...)

### 2018-04-24 submission.Rnw
Tuned text classifier

#check we have captured all the tweets
nrow(df)==(nrow(leave_tweets) + nrow(remain_tweets))
@
It is important to ensure the dataset is balanced, ie there are a similar number of leave and remain tweets. This is indeed the case:
<<>>=
nrow(leave_tweets)
nrow(remain_tweets)
@

@@

In addition to wordclouds, we can plot the most common words for each class of tweet. This is achieve with the function below:
<<>>=
#function to plot wordclouds
plot_wordcount <- function(freq_tab,title_text){
  ggplot(head(freq_tab,20), aes(x=word,y=freq))+
    geom_bar(stat="identity")+
    labs(title=title_text)+
    theme(axis.text.x = element_text(angle = 30, hjust = 1))
}
a<-plot_wordcount(all_tweets$freq_table,"most common words for all tweets")
b<-plot_wordcount(remain$freq_table,"most common words for remain tweets")
c<-plot_wordcount(leave$freq_table,"most common words for leave tweets")
@
The most common word for leave tweeters was 'June', suggesting many of them were focused on reinforcing the timing of the referendum - ensuring people remembered to go out and vote. 'via', 'spread' and 'word' were also common, suggesting attempts to get readers to spread the message. The general theme of many of the common words for leave tweets is of trying to increase virality and therefore increase re-tweets.\\
The most common word for remain tweeters was 'say', followed by 'stay', 'will' and 'bori[s]'. These give less of an impression of a common theme compared to the leave tweets. It is not surprising that 'Boris' is mentioned a lot by the remain tweeters since he proved to be a particularly divisive figure during the campaigning. However, the lack of common theme for leave tweets may point to an overall less coordinated campaign and may have been a contributary factor to the final result of the referendum (this is just conjecture however - it must be stressed that the dataset forms a small minority of all tweets from the campaign period).\\
As a humorous aside 'brexitthemovi[e]' was in the top 20 most common words. I don't remember this term from the campaign period but it is interesting to see such a specific term become so popular.

@@

#checknumber of columns used for model
ncol(as.matrix(dtm))
@
Using the RTextTools package, the tweets are saved in a container-type object. This automatically performs a train-test split. In this case the training set forms 70\% of the data:
<<>>=
container <- create_container(dtm, as.numeric(factor(df$label)),trainSize=1:round(0.7*nrow(df)), testSize=(round(0.7*nrow(df))+1):nrow(df),virgin=FALSE)

@
A number of classification models are trained on the data, and evaluated on the test set. RTextTools automatically creates results and analytics objects which we can use to evaluate model performance:
<<>>=
#train some models and se what we get
models <- train_models(container,algorithms=c("SVM","TREE","BAGGING","BOOSTING","RF"))
results <- classify_models(container, models) #can use for conf matrix
analytics <- create_analytics(container, results)
@

>>>>

#check we have captured all the tweets
nrow(df)==(nrow(leave_tweets) + nrow(remain_tweets))
@
It is important to ensure the dataset is balanced, ie there are a similar number of leave and remain tweets. There are more leave tweets than remain ones, but the difference isn't too great:
<<>>=
nrow(leave_tweets)
nrow(remain_tweets)
@

@@

In addition to wordclouds, we can plot the most common words for each class of tweet. This is achieve with the function below:
<<>>=
#function to plot wordcounts
plot_wordcount <- function(freq_tab,title_text){
  ggplot(head(freq_tab,20), aes(x=word,y=freq))+
    geom_bar(stat="identity")+
    labs(title=title_text)+
    theme(axis.text.x = element_text(angle = 30, hjust = 1))
}
a<-plot_wordcount(all_tweets$freq_table,"most common words for all tweets")
b<-plot_wordcount(remain$freq_table,"most common words for remain tweets")
c<-plot_wordcount(leave$freq_table,"most common words for leave tweets")
@
The most common word for leave tweeters was 'June', suggesting many of them were focused on reinforcing the timing of the referendum - ensuring people remembered to go out and vote. 'via', 'spread' and 'word' were also common, suggesting attempts to get readers to spread the message. The general theme of many of the common words for leave tweets is of trying to increase virality and therefore increase re-tweets.\\
The most common word for remain tweeters was 'say', followed by 'stay', 'will' and 'bori[s]'. These give less of an impression of a common theme compared to the leave tweets. It is not surprising that 'Boris' is mentioned a lot by the remain tweeters since he proved to be a particularly divisive figure during the campaigning. However, the lack of common theme for leave tweets may point to an overall less coordinated campaign and may have been a contributary factor to the final result of the referendum (this is just conjecture however - it must be stressed that the dataset forms a small minority of all tweets from the campaign period).\\
As a humorous aside 'brexitthemovi[e]' was in the top 20 most common words. I don't remember this term from the campaign period but it is interesting to see such a specific term become so popular.

@@

#checknumber of columns used for model
ncol(as.matrix(dtm))
@
Using the RTextTools package, the tweets are saved in a container-type object. This automatically performs a train-test split. In this case the training set forms 70\% of the data:
<<>>=
container <- create_container(dtm, df$label,trainSize=1:round(0.7*nrow(df)), testSize=(round(0.7*nrow(df))+1):nrow(df),virgin=TRUE)

@
A number of classification models are trained on the data, and evaluated on the test set. RTextTools automatically creates results and analytics objects which we can use to evaluate model performance:
<<>>=
#train some models and analyse the results
models <- train_models(container,algorithms=c("SVM","BAGGING","BOOSTING","RF"))
results <- classify_models(container, models) #can use for conf matrix
#analytics <- create_analytics(container, results)
@
Looking at confusion matrices for each of the model types we get the following:
<<>>=
print("SVM")
table(predicted=results$SVM_LABEL,actual=container@testing_codes)
print("Bagging")
table(predicted=results$BAGGING_LABEL,actual=container@testing_codes)
print("Boosting")
table(predicted=results$LOGITBOOST_LABEL,actual=container@testing_codes)
print("Random Forest")
table(predicted=results$FORESTS_LABEL,actual=container@testing_codes)
@
None of these are particularly impressive. Let's take the model with the highest accuracy (the SVM) and tune the hyperparameters. The minimum term frequency is also reduced to incorporate more features into the model training.:
<<>>=
dtm<-make_tdm(df$text,2)
dtm<-dtm$dtm

container <- create_container(dtm, df$label,trainSize=1:round(0.7*nrow(df)), testSize=(round(0.7*nrow(df))+1):nrow(df),virgin=TRUE)

library(e1071)
tuned_svm <- tune.svm(x=container@training_matrix, y=container@training_codes, 
                      validation.x = container@classification_matrix, validation.y=container@testing_codes,
                      gamma=10^(-6:0) , cost=10^(-1:3))

print('model accuracy = ' )
1-tuned_svm$best.performance
@
